---
title: "Statistical Inference Project Part 2"
author: "Aaron Augustine"
date: "July 22, 2015"
output: 
  html_document: 
    keep_md: yes
---
#Background
The purpose of this document is to analyse the ToothGrowth Data in R.  This will include:

1. Load the ToothGrowth data and perform some basic exploratory data analyses, 

2. Provide a basic summary of the data, and 

3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose 

#Analysis
## Load the ToothGrowth data and perform some basic exploratory data analyses
The code and results for this part are given in the Appendix.
```{r, echo=FALSE}
#set working directory
setwd("~/CourseraRClass/StatInf")

#reference necessary libs
library(ggplot2)
```
```{r, echo=FALSE}
data("ToothGrowth")
```

##Provide a basic summary of the data
A great summary of the data can be found at  https://class.coursera.org/statinference-030/forum/thread?thread_id=26.

## Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose
For this section we will conduct 3 hypothesis using non paired T-tests, one for each dose level.  In each case we will test the null hypothesis that the mean length of the odontoblast cells is the same between the form of Orange Juice (OJ) or chemically pure Vitamin C in aqueous solution (VC).  In each case we are assuming unequal variances. 
```{r}
doselist <- unique(ToothGrowth$dose)
for (i in doselist) {
    print(paste("T-Test for Dose Level", i))
    print(t.test(len~supp,data = ToothGrowth, alternative= c("two.sided"), mu=0,
         paired=FALSE, var.equal=FALSE, conf.level= .95, subset=ToothGrowth$dose==i))
}
```
For dose levels .5 and 1 we reject the null hypothesis that the means are equal since they both have p-values less than .05.  For does level 2, we fail to reject the null hypothesis since the p-value is greater than .05.


#Appendix
```{r}
head(ToothGrowth)
str(ToothGrowth)
with(ToothGrowth,table(supp,dose))
```
```{r, fig.height=5, fig.width=5}
qplot(dose,len,data=ToothGrowth, xlab="Dose", ylab="Length", main="", facets=.~supp)
```

